The Way Forward in Parallelizing Dynamic Languages

### 2014-05-02 talk/icooolps2014/position-paper.tex
more random citescites...

\paragraph{HTM} provides us with transactions like any TM system does.
It can be used as a direct replacement for the GIL. However, as is
common with hardware-only solutions, there are quite a few limitations
that can not be lifted easily. For this comparison, we look at the
implementation of Intel in recent Haswell generation CPUs.

HTM in these CPUs works on the level of caches. This has a few
consequences like false-sharing on the cache-line level, and most
importantly it limits the amount of memory that can be accessed within
a transaction. This transaction-length limitation makes it necessary
to have a fallback in place in case this limit is reached. In recent
attempts, the usual fallback is the GIL\cite{odaira14}. In our
experiments, the current generation of HTM proved to be very fragile
and thus needing the fallback very often. Consequently, scalability
suffered a lot from this.

@@

Unfortunately, STM has a big performance problem. One way to approach
this problem is to make STM systems that use the available hardware
better. We are currently working on a STM system that makes use of
several hardware features like virtual memory and memory segmentation.
We further tailor the system to the discussed use case which gives us
an advantage over other STM systems that are more general.  With this
approach, initial results suggest that we can keep the overhead of STM
already below 50\%. A hybrid TM system, which also uses HTM to
accelerate certain tasks, looks like a very promising direction of
research too. In general we believe that further work to reduce the
overhead of STM is very worthwhile.

@@

>>>>

\paragraph{HTM} provides us with transactions like any TM system does.
It can be used as a direct replacement for the GIL\cite{nicholas06,odaira14,fuad10}. However, as is
common with hardware-only solutions, there are quite a few limitations
that can not be lifted easily. For this comparison, we look at the
implementation of Intel in recent Haswell generation CPUs.

HTM in these CPUs works on the level of caches. This has a few
consequences like false-sharing on the cache-line level, and most
importantly it limits the amount of memory that can be accessed within
a transaction. This transaction-length limitation makes it necessary
to have a fallback in place in case this limit is reached. In recent
attempts, the usual fallback is the GIL\cite{odaira14,fuad10}. In our
experiments, the current generation of HTM proved to be very fragile
and thus needing the fallback very often. Consequently, scalability
suffered a lot from this.

@@

Unfortunately, STM has a big performance problem. One way to approach
this problem is to make STM systems that use the available hardware
better. We are currently working on a STM system that makes use of
several hardware features like virtual memory and memory segmentation.
We further tailor the system to the discussed use case which gives us
an advantage over other STM systems that are more general. With this
approach, initial results suggest that we can keep the overhead of STM
below 50\%. A hybrid TM system, which also uses HTM to accelerate
certain tasks, looks like a very promising direction of research
too. In general we believe that further work to reduce the overhead of
STM is very worthwhile.

@@

\bibitem{nicholas06}
  Nicholas Riley and Craig Zilles. 2006. Hardware tansactional memory
  support for lightweight dynamic language evolution. \emph{In
    Companion to the 21st ACM SIGPLAN symposium on Object-oriented
    programming systems, languages, and applications} (OOPSLA
  '06). ACM, New York, NY, USA

\bibitem{fuad10}
  Fuad Tabba. 2010. Adding concurrency in python using a commercial
  processor's hardware transactional memory support. \emph{SIGARCH
  Comput. Archit. News 38}, 5 (April 2010)
